Aspects of RDF Data Constraints in the Social, Behavioural and Economic Sciences

### 2015-04-30 paper/paper7/paper7.tex

\section{Vocabulary Constraint Types}
\label{vocabulary-constraint-types}

The constraint type \emph{vocabulary} guarantees that users do not invent new or use deprecated terms of vocabularies.
%Out-dated classes and properties of previous vocabulary versions can be marked as deprecated.
%The constraint types \emph{context-specific valid classes and properties} (\emph{R-209; R-210}) can be used to specify which classes and properties are valid in which context - here a given vocabulary version.
\emph{Value is valid for datatype} (\emph{R-223}) constraints serve to make sure that all literal values are valid with regard to their datatypes - as stated in the vocabularies.
Thus, it is checked that all date values (e.g., {{\em dcterms:date}) are actually of the datatype \emph{xsd:date} and that \emph{xsd:nonNegativeInteger} values (e.g. \emph{disco:frequency}) are not negative\footnote{For simplicity reasons, we only assign severity levels to \emph{vocabulary constraints} if they differ from \emph{error}.}.
Depending on property datatypes, two different literal values have
a specific ordering with respect to operators like \textless  (\emph{R-43: literal value comparison}).
Start dates (\emph{disco:startDate}), e.g., must be before (\emph{\textless}) end dates (\emph{disco:endDate}).
Consider the following \emph{DL} knowledge base $\mathcal{K}$\footnote{A knowledge base is a collection of formal statements which corresponds to \emph{facts} or what is known explicitly. For simplicity reasons, we do not state namespace prefixes in DL statements.}:
\begin{center}
\begin{DL} 
$\mathcal{K}=\{$ 
	&\ms{isStructuredBy $\sqsubseteq$ $\neg$ column}, \\
	&\ms{TableDescription $\sqcap$ ColumnDescription $\sqsubseteq$ $\perp$}, \\
	&\ms{CategoryStatistics $\equiv$} \\
	&\ms{$\forall$ computationBase.\{valid,invalid\} $\sqcap$ langString}
 \}\\ 
\end{DL}
\end{center}

All properties, not having the same domain and range types, are defined to be pairwise disjoint
(\emph{R-9: disjoint properties}), i.e., no individual \emph{x} can be connected to an individual/literal \emph{y} by disjoint properties (e.g., \emph{phdd:isStructuredBy} and \emph{phdd:column}).
All \emph{PHDD} classes (e.g., \emph{phdd:TableDescription}, \emph{phdd:ColumnDescription}) are pairwise disjoint (\emph{R-7: disjoint classes}),
i.e., individuals cannot be instances of multiple disjoint classes.
It is a common requirement to narrow down the value space of properties by an exhaustive enumeration of valid values (\emph{R-30/37: allowed values}). 
%\emph{Allowed values} (\emph{R-30, R-37}) for properties can be IRIs (matching one or multiple patterns), any literals, allowed literals (e.g. 'red' 'blue' 'green'), and typed literals of one or multiple type(s) (e.g. \emph{xsd:string}). 
\emph{disco:CategoryStatistics}, e.g., can only have \emph{disco:computationBase} relationships to the values \emph{valid} and \emph{invalid} of the datatype \emph{rdf:langString}.
Validation should \emph{exploit sub-super relations} in vocabularies (\emph{R-224}).
If \emph{dcterms:coverage} and one of its sub properties (\emph{dcterms:spatial}, \emph{dcterms:temporal}) are given,
it is checked that \emph{dcterms:coverage} is not redundant with its sub-properties 
which may indicate when the data is verbose/redundant or expressed at a too general level.

\begin{center}
\begin{DL} 
$\mathcal{K}=\{$ &\ms{Variable $\equiv$ $\exists$ concept.Concept}, \\
&\ms{Catalog $\sqsubseteq$ $\forall$ dataset.Dataset}, \\
&\ms{$\exists$ isStructuredBy.$\top$ $\sqsubseteq$ Table}, \\
&\ms{$\top$ $\sqsubseteq$ $\forall$ belongsTo.Concept}
\}\\ 
\end{DL}
\end{center}

We extend $\mathcal{K}$ by \emph{existential quantifications} (\emph{R-86}) enforcing that instances of given classes must have some property relation to individuals/literals of certain types.
Variables, e.g., should have a relation to a theoretical concept ($\mathcal{SL}_{0}$).
%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{Child Care}. 
The default severity level of the constraint is weak, as in most cases research can be continued without having information about the theoretical concept of a variable.

-----

A \emph{universal quantification} (\emph{R-91}) contains all those individuals that are connected by a property only to individuals/literals of particular classes  or data ranges.
%Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets}:
%\begin{DL}
%Catalog $\sqsubseteq$ $\forall$ dataset.Dataset
%\end{DL}
Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets}.
\emph{Property domain} (\emph{R-25, R-26}) and \emph{range} (\emph{R-28, R-35}) constraints restrict domains and ranges of properties.
Only \emph{phdd:Tables}, e.g., can have \emph{phdd:isStructuredBy} relationships and
\emph{xkos:belongsTo} relationships can only point to \emph{skos:Concepts}.
%\emph{Existential quantifications} (\emph{R-86}) enforce that instances of given classes must have some property relation to individuals/literals of certain types.
%Variables, e.g., should have a relation to a theoretical concept ($\mathcal{SL}_{0}$).
%%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
%The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{Child Care}. 
%The default severity level of the constraint is weak, as in most cases research can be continued without having information about the theoretical concept of a variable.
%If a study, e.g., does not contain any data set ($\mathcal{SL}_{2}$), the actual description of the data is missing which may indicate that it is very hard or unlikely to get access to the data.
%If metadata on data files, including the actual data, is missing (especially case and variable quantities; $\mathcal{SL}_{1}$), the description of the data sets and the study is not sufficient.
%Case quantity measures how many cases are collected for a study.
%High case and variable quantities are indicators for high statistical quality and comprehensiveness of the underlying study ($\mathcal{SL}_{1}$).

@@

%\textbf{Validation and Reasoning.}
Some constraint types enable performing reasoning prior to validation which may resolve or cause constraint violations.
With \emph{subsumption} (\emph{R-100}), one can state that \emph{xkos:ClassificationLevel} is a sub-class of \emph{skos:Collection}, i.e., each \emph{xkos:ClassificationLevel} must also be part of the \emph{skos:Collection} class extension.
With \emph{sub properties} (\emph{R-54, R-64}), one can state that \emph{disco:fundedBy} is a sub-property of \emph{dcterms:contributor} - i.e., if a study is funded by an organization, then this organization contributed to this study.
\begin{DL}
ClassificationLevel $\sqsubseteq$ Collection \\
fundedBy $\sqsubseteq$ contributor
\end{DL}

@@

-----

@@

\begin{center}
\begin{DL} 
$\mathcal{K}=\{$ 
  &\ms{fundedBy $\sqsubseteq$ contributor}
 \}\\ 
\end{DL}
\end{center}

By stating that \emph{disco:fundedBy} is a sub property of \emph{dcterms:contributor},
the {\em sub property} (\emph{R-54, R-64}) above assures that if a series is funded by an organization, then the organization must also contribute to the series.
In case the \emph{sub-property} is applied without reasoning and $\mathcal{K}$ contains the triple \ms{disco:fundedBy} \ms{(EU-SILC,} \ms{organization)},
a constraint violation is thrown if $\mathcal{K}$ does not explicitly include the triple \ms{dcterms:contributor} \ms{(EU-SILC,} \ms{organization)}.
If the \emph{sub property} is applied with reasoning, on the other side, the latter triple is derived which resolves the constraint violation.

@@

\section{Simple Constraint Types}

@@

\section{Complex Constraint Types}
\label{complex-constraint-types}

@@

\section{Implementation}
\label{implementation}

@@

\subsection{Evaluation Setup}

>>>>

\subsection{Vocabulary Constraint Types}
\label{vocabulary-constraint-types}

%The constraint type \emph{vocabulary} guarantees that users do not invent new or use deprecated terms of vocabularies.
%Out-dated classes and properties of previous vocabulary versions can be marked as deprecated.
%The constraint types \emph{context-specific valid classes and properties} (\emph{R-209; R-210}) can be used to specify which classes and properties are valid in which context - here a given vocabulary version.
%\emph{Value is valid for datatype} (\emph{R-223}) constraints serve to make sure that all literal values are valid with regard to their datatypes - as stated in the vocabularies.
%Thus, it is checked that all date values (e.g., {{\em dcterms:date}) are actually of the datatype \emph{xsd:date} and that \emph{xsd:nonNegativeInteger} values (e.g. \emph{disco:frequency}) are not negative\footnote{For simplicity reasons, we only assign severity levels to \emph{vocabulary constraints} if they differ from \emph{error}.}.
%Depending on property datatypes, two different literal values have
%a specific ordering with respect to operators like \textless  (\emph{R-43: literal value comparison}).
%Start dates (\emph{disco:startDate}), e.g., must be before (\emph{\textless}) end dates (\emph{disco:endDate}).

%All properties, not having the same domain and range types, are defined to be pairwise disjoint
%(\emph{R-9: disjoint properties}), i.e., no individual \emph{x} can be connected to an individual/literal \emph{y} by disjoint properties (e.g., \emph{phdd:isStructuredBy} and \emph{phdd:column}).
%All \emph{PHDD} classes (e.g., \emph{phdd:TableDescription}, \emph{phdd:ColumnDescription}) are pairwise disjoint (\emph{R-7: disjoint classes}),
%i.e., individuals cannot be instances of multiple disjoint classes.
It is a common requirement to narrow down the value space of properties by an exhaustive enumeration of valid values (\emph{R-30/37: allowed values}). 
%\emph{Allowed values} (\emph{R-30, R-37}) for properties can be IRIs (matching one or multiple patterns), any literals, allowed literals (e.g. 'red' 'blue' 'green'), and typed literals of one or multiple type(s) (e.g. \emph{xsd:string}). 
\emph{disco:CategoryStatistics}, e.g., can only have \emph{disco:computationBase} relationships to the values \emph{valid} and \emph{invalid} of the datatype \emph{rdf:langString}.
%Validation should \emph{exploit sub-super relations} in vocabularies (\emph{R-224}).
%If \emph{dcterms:coverage} and one of its sub properties (\emph{dcterms:spatial}, \emph{dcterms:temporal}) are given,
%it is checked that \emph{dcterms:coverage} is not redundant with its sub-properties 
%which may indicate when the data is verbose/redundant or expressed at a too general level.
Consider the following \emph{DL} knowledge base $\mathcal{K}$\footnote{A knowledge base is a collection of formal statements which corresponds to \emph{facts} or what is known explicitly. For simplicity reasons, we do not state namespace prefixes in DL statements.}:
\begin{center}
\begin{DL} 
$\mathcal{K}=\{$ 
	%&\ms{isStructuredBy $\sqsubseteq$ $\neg$ column}, \\
	%&\ms{TableDescription $\sqcap$ ColumnDescription $\sqsubseteq$ $\perp$}, \\
	&\ms{CategoryStatistics $\equiv$} \\
	&\ms{$\forall$ computationBase.\{valid,invalid\} $\sqcap$ langString}, \\
  &\ms{Variable $\equiv$ $\exists$ concept.Concept}, \\
  &\ms{Catalog $\sqsubseteq$ $\forall$ dataset.Dataset}, \\
  &\ms{$\exists$ isStructuredBy.$\top$ $\sqsubseteq$ Table}, \\
  &\ms{$\top$ $\sqsubseteq$ $\forall$ belongsTo.Concept}
\}\\ 
\end{DL}
\end{center}

%We extend $\mathcal{K}$ by 
\emph{Existential quantifications} (\emph{R-86}) enforce that instances of given classes must have some property relation to individuals/literals of certain types.
Variables, e.g., should have a relation to a theoretical concept (\emph{info}).
%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{Child Care}. 
The default severity level of this constraint is weak, as in most cases research can be continued without having information about the theoretical concept of a variable.
A \emph{universal quantification} (\emph{R-91}) contains all those individuals that are connected by a property only to individuals/literals of particular classes  or data ranges.
%Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets}:
%\begin{DL}
%Catalog $\sqsubseteq$ $\forall$ dataset.Dataset
%\end{DL}
Only \emph{dcat:Catalogs}, e.g., can have \emph{dcat:dataset} relationships to \emph{dcat:Datasets}.
\emph{Property domain} (\emph{R-25, R-26}) and \emph{range} (\emph{R-28, R-35}) constraints restrict domains and ranges of properties:
Only \emph{phdd:Tables}, e.g., can have \emph{phdd:isStructuredBy} relationships and
\emph{xkos:belongsTo} relations can only point to \emph{skos:Concepts}.
%\emph{Existential quantifications} (\emph{R-86}) enforce that instances of given classes must have some property relation to individuals/literals of certain types.
%Variables, e.g., should have a relation to a theoretical concept ($\mathcal{SL}_{0}$).
%%The variables \emph{number of school years} and \emph{highest educational degree}, e.g., are associated with the theoretical concept \emph{education}. 
%The variable \emph{Education at pre-school} is associated with the theoretical concept \emph{Child Care}. 
%The default severity level of the constraint is weak, as in most cases research can be continued without having information about the theoretical concept of a variable.
%If a study, e.g., does not contain any data set ($\mathcal{SL}_{2}$), the actual description of the data is missing which may indicate that it is very hard or unlikely to get access to the data.
%If metadata on data files, including the actual data, is missing (especially case and variable quantities; $\mathcal{SL}_{1}$), the description of the data sets and the study is not sufficient.
%Case quantity measures how many cases are collected for a study.
%High case and variable quantities are indicators for high statistical quality and comprehensiveness of the underlying study ($\mathcal{SL}_{1}$).

@@

%%\textbf{Validation and Reasoning.}
%Some constraint types enable performing reasoning prior to validation which may resolve or cause constraint violations.
%With \emph{subsumption} (\emph{R-100}), one can state that \emph{xkos:ClassificationLevel} is a sub-class of \emph{skos:Collection}, i.e., each \emph{xkos:ClassificationLevel} must also be part of the \emph{skos:Collection} class extension.
%With \emph{sub properties} (\emph{R-54, R-64}), one can state that \emph{disco:fundedBy} is a sub-property of \emph{dcterms:contributor} - i.e., if a study is funded by an organization, then this organization contributed to this study.
%\begin{DL}
%ClassificationLevel $\sqsubseteq$ Collection \\
%fundedBy $\sqsubseteq$ contributor
%\end{DL}

@@

@@

%\begin{center}
%\begin{DL} 
%$\mathcal{K}=\{$ 
  %&\ms{fundedBy $\sqsubseteq$ contributor}
 %\}\\ 
%\end{DL}
%\end{center}
%
%By stating that \emph{disco:fundedBy} is a sub property of \emph{dcterms:contributor},
%the {\em sub property} (\emph{R-54, R-64}) above assures that if a series is funded by an organization, then the organization must also contribute to the series.
%In case the \emph{sub-property} is applied without reasoning and $\mathcal{K}$ contains the triple \ms{disco:fundedBy} \ms{(EU-SILC,} \ms{organization)},
%a constraint violation is thrown if $\mathcal{K}$ does not explicitly include the triple \ms{dcterms:contributor} \ms{(EU-SILC,} \ms{organization)}.
%If the \emph{sub property} is applied with reasoning, on the other side, the latter triple is derived which resolves the constraint violation.

@@

\subsection{Simple Constraint Types}

@@

\subsection{Complex Constraint Types}
\label{complex-constraint-types}

@@

\subsection{Experimental Setup}
\label{implementation}

@@

\ke{Überarbeiten und mit nächstem teil mergen. Nur das erzählen, was für die Auswertung auf den Datensets relevant ist. 114 constraints, 3 vokabulare, wie wurden die datasets daraufhin validiert.}
